---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
The data is already located in the right location. First, let's load it into a data frame, and run `str` and `summary` on the data frame to get an idea of what's in it:
```{r}
allDF <- read.csv("activity.csv", header = TRUE)
str(allDF)
summary(allDF)
head(allDF)
# in case we need to modify the date field:
# myData$Date <- as.Date(myData$Date, format = "%d/%m/%Y")
```
For the next section we are interested in daily totals, so let's summarise these and we'll take a look at the result by running `head`:
```{r, message=F, warning=F }
library("dplyr")
```
```{r}
dailyTotals <- allDF %>%
  group_by(date) %>%
  summarise(daily = sum(steps))

head(dailyTotals)
```
## What is mean total number of steps taken per day?
```{r}
summary(dailyTotals$daily)
hist(dailyTotals$daily, xlab = "Steps", breaks = 5)
meanDaily <- mean(dailyTotals$daily, na.rm = TRUE)
meanDaily
abline(v = meanDaily, col = "blue", lwd = 2)
```

## What is the average daily activity pattern?
```{r}
dailyMean <- allDF %>%
        group_by(interval) %>%
        summarise(average = mean(steps, na.rm = TRUE))

head(dailyMean)
tail(dailyMean)

with(dailyMean, {
        plot(x = interval,
             y = average,
             type = "l",
             xlab = "Interval",
             ylab = "Steps"
        )
} )
```


## Imputing missing values
```{r}
imputedDF <- allDF
sum(is.na(imputedDF$steps))
for(i in seq(from = 0, to = 2355, by = 5)) {
        imputedDF$steps[is.na(imputedDF$steps) & imputedDF$interval == i] <- 0
        #if(imputedDF[is.na(imputedDF$steps),]
}
i
sum(is.na(imputedDF$steps))
```


## Are there differences in activity patterns between weekdays and weekends?
